---
title: "p8105_hw1_dm3175"
author: "Devon Morgan"
date: "9/20/2018"
output: github_document
---

# Background

This file contains the solutions to Homework 1 assignment for Data Science I (P8105). The due date for Homework 1 is 9/21/2018.

# Problem 1

The goal of problem 1 is to explore variable types and coercion. 

## Part 1: Naming Variables and Creation of Data Frame

The following variables and corresponding interpretation will be considered for Problem 1. For this example, fruit types and condition ("good" or "bad") were used:

Variable Name  |  Definition
---------------|---------------------------------------------------------
**uniform_dist**| A random sample of size 10 from uniform [0,5] distribution
**greater_than_two** | A logical vector indicating whether elements of the sample are greater than 2
**fruit_type** | A length-10 character vector for fruit type
**fruit_condition** | A length-10 factor vector for fruit condition: "Bad" or "Good"

Next, a **data frame** was created containing the variables listed above:

```{r setup problem 1, include = FALSE}
# Load relevant package
library(tidyverse)

```

```{r create data frame problem 1}
# Seed is set to ensure reproducibility 
set.seed(1234)

# Create data frame 
problem_1_df = tibble (
  uniform_dist = runif(10, min = 0, max = 5),
  greater_than_two = uniform_dist > 2,
  fruit_type = c("Apple","Orange","Grape","Pear","Kiwi","Banana","Berry","Cherry","Peach","Mango"),
  fruit_condition = factor(c("Good","Bad","Bad","Good","Bad","Good","Good","Good","Good","Bad"))
)

```

## Part 2: Mean of All Variables 

Next, an attempt at taking the **mean** of each variable was executed: 

```{r mean of all variables}
mean(problem_1_df$uniform_dist)
mean(problem_1_df$greater_than_two)
mean(problem_1_df$fruit_type)
mean(problem_1_df$fruit_condition)

```

The mean can be successfully taken on both the `uniform_dist` variable because it is numerical and the `greater_than_two` variable because it is a logical variable. The logical variable codes "TRUE" responses as 1 and "FALSE" responses as 0, which allows the mean to be interpreted as the proportion of "TRUE" values in the dataset. The mean cannot be taken of `fruit_type` or `fruit_condition` because these variables are character and factor variables, respectively, which do not allow for a numeric mean to be computed (note the error message produced).

## Part 3: Conversion of variables to different types

Next, the logical, character, and factor variables were converted to numerical variables using the ` as.numeric` function. The `as.numeric` function works successfully to convert logical (`greater_than_two`) and factor (`fruit_condition`) variables to numerical form. The function produces an error when applied directly to a character (`fruit_type`) variable because the code needs further specification on how to code the character values as numbers (i.e., it does not automatically know what numerical value to assign to what characters). 

```{r converting variable types, results = 'hide', warning=FALSE}
as.numeric(problem_1_df$greater_than_two)
as.numeric(problem_1_df$fruit_type)
as.numeric(problem_1_df$fruit_condition)

```

To convert the character variable into a numerical variable, the character variable (`fruit_type`) was first converted into a factor variable and then numerical. Similarly, the factor variable (`fruit_condition`) was converted into a character variable and then numeric. 

```{r factor and character variable conversions, warning=FALSE}
# Conversion of "fruit_type" to factor then numerical
factor_fruit_type = as.factor(problem_1_df$fruit_type)
numeric_fruit_type = as.numeric(factor_fruit_type)

# Conversion of "fruit_condition" to character then numerical
char_fruit_condition = as.character(problem_1_df$fruit_condition)
numeric_fruit_condition = as.numeric(char_fruit_condition)

```

The conversion of character variable to factor variable and then numerical variable works successfully, as there is an intermediate step converting to a factor variable prior to numerical. Changing the order from factor variable to character to numeric does not work (similar to above, the code needs further specification on how to code the character variable into numerical). **In sum, the order of converting variables matters.** 

# Problem 2

The goal of problem 2 is to plot and use inline R code. 

## Part 1: Naming Variables and Creation of Data Frame

The following variables and corresponding interpretation will be considered for problem 2:


Variable Name  |  Definition
---------------|---------------------------------------------------------
**x**         | a random sample of size 1000 from a standard normal distribution
**y**         | a random sample of size 1000 from a standard normal distribution
**vec_logical** | a logical vector indicating whether x + y > 0
**vec_numerical** | a numeric vector created by coercing the logical vector "vec_logical"
**vec_factor** | a factor vector created by coercing the logical vector "vec_logical"

Next, a **data frame** was created containing the variables listed above:

```{r setup problem 2, include = FALSE}
# Load relevant package
library(tidyverse)

```

```{r create data frame problem 2}
# Seed is set to ensure reproducibility 
set.seed(12345)

# Create data frame 
problem_2_df = tibble (
  x = rnorm(1000),
  y = rnorm(1000),
  vec_logical = x + y > 0,
  vec_numerical = as.numeric(vec_logical),
  vec_factor = as.factor(vec_logical)
)

```

The data frame contains `r nrow(problem_2_df)` rows and `r ncol(problem_2_df)` columns. The mean of `x` is `r mean(problem_2_df$x)` and the median of `x` is `r median(problem_2_df$x)`. The proportion of cases for which the logical vector `vec_logical` is TRUE is `r mean(problem_2_df$vec_logical)`.

## Part 2: Scatterplots of Data Frame

A scatterplot of `y` vs `x` was created that color points using `vec_logical`. This scatter plot was saved into the project directory. 

```{r scatterplot of Y vs. X with logical vector colored}
ggplot(problem_2_df, aes(x = x, y = y)) + geom_point(aes(color = vec_logical))

ggsave("Problem_2_Scatterplot.pdf", height = 4, width = 6)

```

A scatterplot of `y` vs `x` was created that color points using `vec_numerical`. The color scale on this scatterplot demonstrates that numerical vectors can be colored according to a gradient scale.

```{r scatterplot of Y vs. X with numeric vector colored}
ggplot(problem_2_df, aes(x = x, y = y)) + geom_point(aes(color = vec_numerical))

```

A scatterplot of `y` vs `x` was created that color points using `vec_factor`. The color scale on this scatterplot is similar to that of the logical vector with two possible colors, showing that factor variables color similar to logical vectors.

```{r scatterplot of Y vs. X with factor vector colored}
ggplot(problem_2_df, aes(x = x, y = y)) + geom_point(aes(color = vec_factor))

```

