---
title: "p8105_hw1_dm3175"
author: "Devon Morgan"
date: "9/17/2018"
output: github_document
---

# Background

This file contains the solutions to Homework 1 assignment for Data Science I (P8105). The due date for Homework 1 is 9/21/2018.

# Problem 1

The goal of problem 1 is to explore variable types and coercion. 

## Part 1: Naming Variables and Creation of Data Frame

The following variables and corresponding interpretation will be considered for Problem 1. For this example, fruit types and condition (Good or Bad) were used:

Variable Name  |  Definition
---------------|---------------------------------------------------------
**uniform_dist**| A random sample of size 10 from uniform [0,5] distribution
**greater_than_two** | A logical vector indicating whether elements of the sample are greater than 2
**fruit_type** | A length-10 character vector for fruit type
**fruit_condition** | A length-10 factor vector for fruit condition: "Bad" or "Good"

Next, a **data frame** was created containing the variables listed above:

```{r include = FALSE}
# Load relevant package

library(tidyverse)

```

```{r}
# Seed is set to ensure reproducibility 
set.seed(1234)

# Create data frame 
problem_1_df = tibble (
  uniform_dist = runif(10, min = 0, max = 5),
  greater_than_two = uniform_dist > 2,
  fruit_type = c("Apple","Orange","Grape","Pear","Kiwi","Banana","Berry","Cherry","Peach","Mango"),
  fruit_condition = factor(c("Good","Bad","Bad","Good","Bad","Good","Good","Good","Good","Bad"))
)

```


## Part 2: Conversion of variables to different types

# Problem 2

