---
title: "p8105_hw1_dm3175"
author: "Devon Morgan"
date: "9/17/2018"
output: github_document
---

# Background

This file contains the solutions to Homework 1 assignment for Data Science I (P8105). The due date for Homework 1 is 9/21/2018.

# Problem 1

The goal of problem 1 is to explore variable types and coercion. 

## Part 1: Naming Variables and Creation of Data Frame

The following variables and corresponding interpretation will be considered for Problem 1. For this example, fruit types and condition (Good or Bad) were used:

Variable Name  |  Definition
---------------|---------------------------------------------------------
**uniform_dist**| A random sample of size 10 from uniform [0,5] distribution
**greater_than_two** | A logical vector indicating whether elements of the sample are greater than 2
**fruit_type** | A length-10 character vector for fruit type
**fruit_condition** | A length-10 factor vector for fruit condition: "Bad" or "Good"

Next, a **data frame** was created containing the variables listed above:

```{r Setup, include = FALSE}
# Load relevant package

library(tidyverse)

```

```{r Create Data Frame}
# Seed is set to ensure reproducibility 
set.seed(1234)

# Create data frame 
problem_1_df = tibble (
  uniform_dist = runif(10, min = 0, max = 5),
  greater_than_two = uniform_dist > 2,
  fruit_type = c("Apple","Orange","Grape","Pear","Kiwi","Banana","Berry","Cherry","Peach","Mango"),
  fruit_condition = factor(c("Good","Bad","Bad","Good","Bad","Good","Good","Good","Good","Bad"))
)

```

Next, an attempt at taking the **mean** of each variable was executed: 

```{r Mean of All Variables}
mean(problem_1_df$uniform_dist)
mean(problem_1_df$greater_than_two)
mean(problem_1_df$fruit_type)
mean(problem_1_df$fruit_condition)

```

The mean can be successfully taken on both the "uniform_dist" variable because it is numerical and the "greater_than_two" variable because it is a logical variable. The logical variable codes "TRUE" responses as 1 and "FALSE" responses as 0, which allows the mean to be interpreted as the proportion of "TRUE" values in the dataset. The mean cannot be taken of "fruit_type" or "fruit_condition" because these variables are character and factor variables, respectively, which do not allow for a numeric mean to be computed. 

## Part 2: Conversion of variables to different types

Next, the logical, character, and factor variables were converted to numerical variables using the ` as.numeric` function. The `as.numeric` function works successfully to convert logical ("greater_than_two") and factor ("fruit_condition") variables to numerical form. The function produces an error when applied directly to a character ("fruit_type") variable because the code needs further specification on how to code the character values as numbers. 

```{r Converting Variable Types, results = 'hide', warning=FALSE}
as.numeric(problem_1_df$greater_than_two)
as.numeric(problem_1_df$fruit_type)
as.numeric(problem_1_df$fruit_condition)

```

To convert the character variable into a numerical variable, the character variable ("fruit_type") was first converted into a factor variable and then numerical. Similarly, the factor variable ("fruit_condition") was converted into a character

```{r Factor and Character Variable Conversions, warning=FALSE}
# Conversion of "fruit_type" to factor then numerical
factor_fruit_type = as.factor(problem_1_df$fruit_type)
numeric_fruit_type = as.numeric(factor_fruit_type)

# Conversion of "fruit_condition" to character then numerical
char_fruit_condition = as.character(problem_1_df$fruit_condition)
numeric_fruit_condition = as.numeric(char_fruit_condition)

```

# Problem 2

